---
title: "P8105 Homework 5"
author: "Derek Lamb"
date: "`r Sys.Date()`"
output: github_document
---
```{r load packages, message = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1
I imported the data, noting that missing data are coded as `Unknown`. I also corrected a typo, where a homicide in Tulsa, OK was misclassified as in AL, and corrected the state representations (Wisconsin was `wi`). I then summarized the homicide data by city and state.
```{r data import}
df_homicide <- read_csv("data/homicide-data.csv", 
                        na = "Unknown") |> 
  mutate(
    state = case_when(
      city == "Tulsa" ~ "OK",
      city != "Tulsa" ~ state),
    state = str_to_upper(state),
    city_state = str_c(city, ", ", state)
    ) |> 
  group_by(city_state) |> 
  summarize(n_homicide = n(),
            n_unsolved = sum(disposition %in% c("
                                                Closed without arrest",
                                                "Open/No arrest")))
```

Within Baltimore, MD, I calculated the proportion of unsolved cases and a confidence interval for this value.
```{r baltimore prototype}
df_homicide |> 
  filter(city_state == "Baltimore, MD") |> 
  mutate(df = broom::tidy(prop.test(x = n_unsolved, 
                                    n = n_homicide,
                                    p = 0.5,
                                    alternative = "two.sided")))
```

To improve workflow, I wrote a function to perform a z-test and construct a confidence interval under normal approximations of the data.
```{r define homicide function}
z_homicide = function(x, n){
  prop.test(x = x, n = n,
            p = 0.5, alternative = "two.sided") |> 
  broom::tidy()
}
```

Now I use this function to create a plot of the confidence interval for every city.
```{r create prop plot}
df_homicide |> 
  mutate(z_out = map2(n_unsolved, n_homicide, z_homicide)) |> 
  unnest(z_out) |> 
  select(city_state, estimate, conf.low, conf.high) |> 
  mutate(city_state = fct_reorder(city_state, estimate)) |> 
  ggplot(aes(x = estimate, y = city_state)) + 
  geom_point() + 
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high)) + 
  labs(
      title = "Point and Interval Estimates of Homicide Closure",
      x = "Proportion of Unsolved Homicides",
      y = "City & State"
  )
```

# Problem 2
